Glyph-maps for visualizing climate data and models

### 2011-09-23 paper.tex
caption edits

\subsection{Irregular locations}
Figure \ref{fig:irregular} shows linear trends for monthly average surface temperature at stations in the USHCN (Version 2) ground station network.  It illustrates three challenges in working with irregular locations: there is no natural choice for icon size, icons can potentially overlap, and comparisons rely more heavily on the reference guides.  The icon sizes were chosen manually for these two plots, although the icons are displayed at about the same size, they correspond to 1$^o$ and 0.5$^o$ square respectively in the geographical coordinates. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.55\linewidth]{usa-lin-overlap}%
  \includegraphics[width=0.35\linewidth]{ghcn-mountains}%
  \caption{Glyph-maps of the linear trend in ground station temperature values, 1950-2010, of all continental US USHCN stations (left) and stations in the mountain states (right).} 
  \label{fig:irregular}
\end{figure}
The two panels illustrate one approach to the overlapping problem: zooming.  There is an interplay between icon size (in geographical units), plotting area (in geographical units) and plot size (in display units).  Icons need to be big enough to be readable on the display, but small enough to avoid overlapping.  Readability can be maintained and overlapping reduced by decreasing the icon size and simultaneously increasing the display size or decreasing the plotting area (as done in the right panel in Figure \ref{fig:irregular}).  However, there are generally limitations in the maximum display size (or resolution) and the utility of plotting very small areas.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{usa-season-collapsed}%
  \includegraphics[width=0.5\linewidth]{usa-lin-collapse}%
  \caption{Glyph-maps of ground station temperature values, with nearby locations collapsed.1950-2010.}
  \label{fig:irregular-collapsed}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{usa-season-grid}%
  \includegraphics[width=0.5\linewidth]{usa-lin-grid}%
  \caption{Glyph-maps of ground station temperature values, with locations rounded to 1$^o$. 1950-2010.}
  \label{fig:irregular-grid}
\end{figure}

@@

>>>>

\subsection{Irregular locations}
Figure \ref{fig:irregular} shows linear trends for monthly average surface temperature at stations in the USHCN (Version 2) ground station network.  It illustrates three challenges in working with irregular locations: there is no natural choice for icon size, icons can potentially overlap, and comparisons rely more heavily on the reference guides.  The icon sizes were chosen manually for these two plots, although the icons are displayed at about the same size, they correspond to 1$^o$ and 0.5$^o$ square respectively in the geographical coordinates. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.55\linewidth]{usa-lin-overlap}%
  \includegraphics[width=0.35\linewidth]{ghcn-mountains}%
  \caption{Glyph-maps of the linear trend in ground station temperature values, 1950--2010, for all continental US USHCN stations (left) and stations in the mountain states (right).} 
  \label{fig:irregular}
\end{figure}
The two panels illustrate one approach to the overlapping problem: zooming.  There is an interplay between icon size (in geographical units), plotting area (in geographical units) and plot size (in display units).  Icons need to be big enough to be readable on the display, but small enough to avoid overlapping.  Readability can be maintained and overlapping reduced by decreasing the icon size and simultaneously increasing the display size or decreasing the plotting area (as done in the right panel in Figure \ref{fig:irregular}).  However, there are generally limitations in the maximum display size (or resolution) and the utility of plotting very small areas.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{usa-season-collapsed}%
  \includegraphics[width=0.5\linewidth]{usa-lin-collapse}%
  \caption{Glyph-maps of seasonal patterns (left) and linear trends (right) of USHCN stations, 1950--2010.  Nearby stations whose icons would overlap have been collapsed.  Each location has a glpyh in the seasonal plot and colour of the tiles corresponds to the average temperature of the locations (red=warm, blue=cool).  Heavy lines in the linear trend plot correspond to average trends over more than one location.}
  \label{fig:irregular-collapsed}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{usa-season-grid}%
  \includegraphics[width=0.5\linewidth]{usa-lin-grid}%
  \caption{Glyph-maps of seasonal patterns (left) and linear trends (right) of USHCN stations, 1950--2010.  Station locations have been rounded to the nearest degree Each location has a glpyh in the seasonal plot and colour of the tiles corresponds to the average temperature of the locations (red=warm, blue=cool).  Heavy lines in the linear trend plot correspond to average trends over more than one location.}
  \label{fig:irregular-grid}
\end{figure}

@@

USHCN data provided by NOAA NCDC from their website \url{http://cdiac.ornl.gov/ftp/ushcn_v2_monthly/}.
